---
title: "Awards"
author: "Yang"
date: "February 11, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
awards<-read.csv("awards.csv")
attach(awards)
```

**Comments on significance of predictors**

##Model 1: Math as a continuous predictor

###Fit 
```{r}
m1=glm(numawards~1+math,family = poisson(link = log),x=TRUE)
summary(m1)
```

###Comment on the significance of math as a predictor
Math is a significant predictor for number of awards since the p-value is small(<0.05).

##Model 2: Prog as a factor predictor
###Fit
```{r}
m2=glm(numawards~1+as.factor(prog),family=poisson(link=log),x=TRUE)
summary(m2)
```

###Comments on significance of params
The model has 3 parameters. And $\beta_0~,\beta_1$ are statistically important since they have small p-value(1.38e-06 and 3.59e-06 respectively) compared with 0.05 while $\beta_2$ is not statistically significant in our model since it has a large p-value 0.679.\newline

###Wald test
```{r}
I = t(m2$x)%*%diag(m2$weights)%*%m2$x
I_inv = solve(I)
sd <- sqrt(diag(I_inv))
p_value <- pchisq((m2$coefficients/sd)^2,df=2,lower.tail=FALSE)
p_value
```
Meanwhile, according to the wald test, factor2("Academic") is statistically significant with p-value 2.174618e-05 while factor3("Vocational") is not statistically significant due to its p-value 9.180740e-01. 

###Likelihood ratio test
```{r LR test}
library(lmtest)
m_null=glm(numawards~1,family=poisson(link=log),x=TRUE)
test=lrtest(m2,m_null)
test
test$`Pr(>Chisq)`
```
According to the likelihood ratio test resulting in p-value=2.786791e-12<0.05, we reject the null model and can conclude that prog is a significant predictor.



##Model 3: numawards~1+math+as.factor(prog)

###Fit
```{r model3:numawards~1+math+as.factor(prog)}
m3=glm(numawards~1+math+as.factor(prog), family=poisson(link='log'),x=TRUE)
summary(m3)
```
model 3:numawards~1+math+as.factor(prog), has 4 parameters.\newline

###Interpretation 
Expected number of awards should increase by $e^{0.07015}$ if **math** is increased by 1.\newline 
The difference in expected number of awards between student enrolled in **academic** program and in general program is $e^{1.08386}$.\newline
The difference in expected number of awards between student enrolled in **vocational** program and in general program is $e^{0.36981}$.\newline
\newline

##Model 4: numawards~1 + math * as.factor(prog)

###Fit
```{r model4:1 + math * as.factor(prog)}
m4=glm(numawards~1+math*as.factor(prog), family=poisson(link='log'),x=TRUE)
summary(m4)
```
model 4: numawards ~ 1 + math * as.factor(prog), has 6 parameters.\newline

###Interpretation
Expected number of awards should increase by $e^{0.044}$ if **math** is increased by 1.\newline
The difference in expected number of awards between student enrolled in **academic** program and in general program is $e^{-0.44107}$.\newline
The difference in expected number of awards between student enrolled in **vocational** program and in general program is $e^{-0.84473}$.\newline
The difference in expected number of awards between student enrolled in **academic** program and in general program is expected to increase by $e^{0.02841}$ if **math** increase by 1.\newline
The difference in expected number of awards between student enrolled in vocational program and in general program is expected to increase by $e^{0.02841}$ if **math** increase by 1.\newline

##Using Plot to Analyze

###model1
```{r}
awards$m1=m1$fitted.values
awards$m2=m2$fitted.values
awards$m3=m3$fitted.values
awards$m4=m4$fitted.values
X=seq(30,70,5)
df= split(awards, cut(awards$math, breaks = seq(30,75,5)))
Mean_award <- function(x){
  mean(x$numawards)
}
Mean_prediction_m1 <- function(x){
  mean(x$m1)
}
True_awards = sapply(df, Mean_award)
Estimated_awards_1=sapply(df,Mean_prediction_m1)

plot(X,True_awards,main="Fitted line")
lines(X,Estimated_awards_1)
plot(m1)
```

###model2
```{r}
Mean_prediction_m2 <- function(x){
  mean(x$m2)
}
Estimated_awards_2=sapply(df,Mean_prediction_m2)
plot(as.factor(awards$prog),awards$numawards)
academic=subset(awards,prog==2)
vocational=subset(awards,prog==3)
general=subset(awards,prog==1)
points(awards$prog,awards$numawards)
plot(m2)
```

###model3
*The black dots are true datapoints and the red dots are # of awards estimated by model 3.*
```{r}
plot(academic$math,academic$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Students Enrolled in Academic")
points(academic$math,academic$m3,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))

plot(vocational$math,vocational$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Students Enrolled in vocational")
points(vocational$math,vocational$m3,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))

plot(general$math,general$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Students Enrolled in General")
points(general$math,general$m3,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))


plot(awards$math,awards$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Math against # of rewards among three categories")
points(awards$math,awards$m3,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))
plot(m3)
```

###model4
**Note: The black dots are true datapoints and the red dots are # of awards estimated by model 4.**
```{r}
plot(academic$math,academic$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Students Enrolled in Academic")
points(academic$math,academic$m4,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))

plot(vocational$math,vocational$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Students Enrolled in vocational")
points(vocational$math,vocational$m4,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))

plot(general$math,general$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Students Enrolled in General")
points(general$math,general$m4,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))


plot(awards$math,awards$numawards,col="black",pch=20,xlab="math",ylab = "# of awards",main = "Math against # of rewards among three categories")
points(awards$math,awards$m4,col="red")
legend(x=3.5,y=14, legend=c("True", "Predicted"),fill=c("black", "red"))
plot(m4)
```


##Compare Deviance
```{r Compare Deviance}
m1$deviance
m2$deviance
m3$deviance
m4$deviance
```

Based on the deviance of model 1 to 4, we can conclude that model 4 is the best with smallest deviance 189.1016.



